---
title: "Compare factors & factors after regressing out GTEx PEER factors"
author: "XSun"
date: "2021-03-12"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## **Introduction**

In earlier [PC results](https://xsun1229.github.io/factor_analysis/kegg_pcs_factor_traits.html), we found many associations are driven by PC4/PC5. We guess that the top few PCs may largely capture the batch/technical effects, and may not be biologically interesting. So we regress out the 60 PEER factors provided by GTEx and use the residuals to do the association test again.

## **Results comparison**

This table shows the pc distribution for all pairs. "bf" here means "before regressing out PEER factors", "af" here means "after regressing out PEER factors"

```{r echo=FALSE}
load("/project2/xinhe/xsun/website/factor_analysis/output/kegg_pc_sum.rdata")
pc_sum1 <- pc_sum
colnames(pc_sum1)[2:3] <- c("n_bf","percent_bf")
load("/project2/xinhe/xsun/website/factor_analysis/output/kegg_pc_sum_reg.rdata")
colnames(pc_sum)[2:3] <- c("n_af","percent_af")

sum <- merge(pc_sum1,pc_sum, by = "pc")

DT::datatable(sum, options = list(pageLength =20))
```


This table shows the pc distribution for significant pairs (p_values < 0.05 in resampling).


```{r echo=FALSE}
load("/project2/xinhe/xsun/website/factor_analysis/output/kegg_pc_sum_top.rdata")
pc_sum1 <- pc_sum
colnames(pc_sum1)[2:3] <- c("n_bf","percent_bf")
load("/project2/xinhe/xsun/website/factor_analysis/output/kegg_pc_sum_top_reg.rdata")
colnames(pc_sum)[2:3] <- c("n_af","percent_af")

sum <- merge(pc_sum1,pc_sum, by = "pc")

DT::datatable(sum, options = list(pageLength =20))
```
