---
title: "Effect Size Consistency Tests (SNP FDR < 0.2, FDR corrected)"
author: "XSun"
date: "2022-10-15"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r echo=FALSE, message=FALSE, warning=FALSE}
suppressMessages(library(ggplot2))
suppressMessages(library(gridExtra))
suppressMessages(library(tidyr))
suppressMessages(library(forcats))
```

# Introduction

Workflow:

1. Association test: factor ~ snp + covariates 

2. For the pairs have multiple snps passing FDR < 0.2, do ECT

# Results

## The number of SNPs passing FDR < 0.2 in association tests

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE,fig.height=10, fig.width=15}

type <- c("B cell (CD19+)","Monocyte (CD14+)","Leukocyte (CD15+)","Platelet","T cell (CD4+)","T cell (CD8+)")
path <- c("4.B_cell_kegg","6.CD14_positive_monocyte_kegg","8.CD15_positive_leukocyte_kegg","16.platelet_kegg","10.T_cell_kegg","14.thymocyte_kegg")


p <- list()
for (t in 1:length(type)) {
  
  folder_path <- paste0("/project2/xinhe/xsun/CEDAR/",path[t],"/9.pcs_40genes_rerun/2.ECT/")

  load(paste0(folder_path,"results/number_snp_fdr02_allpairs.rdata"))
  
  count_vector <- c()
  for (i in 1:length(snp_num_fdr02)) {
    count_vector <- c(count_vector,snp_num_fdr02[[i]])
  }
  
  summary <- as.data.frame(table(count_vector))
  
  p[[t]] <- ggplot(summary, aes(x = count_vector, y =Freq)) +
    geom_bar(stat = "identity",fill = "steelblue") +
    
    theme_bw(base_line_size =0.3) +
    
    ggtitle(type[t]) + theme(plot.title = element_text(hjust = 0.5)) +
    labs(x = "Number of SNPs pass FDR < 0.2 for each pair",
         y = "Count") +
    
    geom_text(aes(label = Freq), colour = "black",size =2,vjust = -0.1) +
    
    theme(axis.title.x = element_text(size = 10),
          axis.text.x = element_text(size = 8, color = "black"),
          axis.title.y = element_text(size = 10),
          axis.text.y = element_text(size = 8, color = "black"),
          text= element_text(family="Times New Roman")) 
  
}

all <- grid.arrange(p[[1]],p[[2]],p[[3]],p[[4]],p[[5]],p[[6]], nrow = 2)


```

## ECT results

### Number of pairs passing different ECT FDR cutoff 

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE,fig.height=5, fig.width=5}

path <- c("4.B_cell_kegg","6.CD14_positive_monocyte_kegg","8.CD15_positive_leukocyte_kegg","16.platelet_kegg","10.T_cell_kegg","14.thymocyte_kegg")

x <- c()
y <- c()
for (i in 1:length(path)){
  folder_path <- paste0("/project2/xinhe/xsun/CEDAR/",path[i],"/9.pcs_40genes_rerun/2.ECT/")
  
  load(paste0(folder_path,"results/ECT_summary_table.rdata"))
  x[i] <- sum(df_sum$fdr_ECT < 0.1)
  y[i] <- sum(df_sum$fdr_ECT < 0.2) 
  
  
}

type <- c("CD19+","CD14+","CD15+","PLT","CD4+","CD8+")


data <- data.frame(x, y,type)
df <- gather(data, event, total, x:y)


bar <- ggplot(df, aes(type, total, fill=event)) +
  
  theme_bw(base_line_size =0.3) +

  geom_bar(stat = "identity", position = 'dodge', alpha = 0.2) +
  scale_fill_manual(name="group", values=c("red", "blue"),labels=c("FDR < 0.1","FDR < 0.2")) +
  
  aes(x = fct_inorder(type)) + 
  
  labs(x = "Cell types", y = "Count") +
  
  geom_text(aes(label = total),position = position_dodge(0.9),
            color="black",size = 3,vjust = -0.3) +
  
  ylim(0,80) +
  
  theme(axis.title.x = element_text(size = 12),
        axis.text.x = element_text(size = 10, color = "black"),
        axis.title.y = element_text(size = 12),
        axis.text.y = element_text(size = 10, color = "black"),
        text= element_text(family="Times New Roman"))

bar

```

### ECT qqplots and rsq barplots for all pairs have >=3SNPs

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE,fig.height=10, fig.width=15}


type <- c("B cell (CD19+)","Monocyte (CD14+)","Leukocyte (CD15+)","Platelet","T cell (CD4+)","T cell (CD8+)")
path <- c("4.B_cell_kegg","6.CD14_positive_monocyte_kegg","8.CD15_positive_leukocyte_kegg","16.platelet_kegg","10.T_cell_kegg","14.thymocyte_kegg")

p1 <- list()
p2 <- list()
for (i in 1:length(type)){ 
  
  file_ect_results <- paste0("/project2/xinhe/xsun/CEDAR/",path[i],"/9.pcs_40genes_rerun/2.ECT/results/ECT_results.rdata")
  load(file_ect_results)
  
  file_ect_summary <- paste0("/project2/xinhe/xsun/CEDAR/",path[i],"/9.pcs_40genes_rerun/2.ECT/results/ECT_summary_table.rdata")
  load(file_ect_summary)
  
  plot_sum <- df_sum[df_sum$num_snp_pass_cutoff >= 3,] ######
  
  plotdata <- data.frame(observed = -log10(sort(as.numeric(as.character(plot_sum$pval_ECT)))),
                         expected = -log10(ppoints(nrow(plot_sum))))
  #  
  
  p1[[i]] <- ggplot(plotdata) + theme_bw(base_line_size =0.3,) +
    geom_point(aes(expected, observed), shape = 1, size = 3, color= "steelblue") +
    geom_abline(intercept = 0, slope = 1, alpha = 0.6, color = "red") +
    ggtitle(type[i]) + theme(plot.title = element_text(hjust = 0.5)) +
    labs(x = expression(paste("Expected -log"[10],"(p-value)")),
         y = expression(paste("Observed -log"[10],"(p-value)"))) +
    
    theme(axis.title.x = element_text(size = 22),
          axis.text.x = element_text(size = 20, color = "black"),
          axis.title.y = element_text(size = 22),
          axis.text.y = element_text(size = 20, color = "black"),
          text= element_text(family="Times New Roman")) 
  
  
  #### rsq barplot (selected data)
  rsq_resampling_all <- c()
  rsq_fitting_all <- c()
  for (t in 1:nrow(plot_sum)) {
    
    trait <- plot_sum$trait[t]
    factor <- plot_sum$factor[t]
    
    r_rsp_pair <- r_rsp_all[[trait]][[factor]]
    set.seed(100)
    select <- sample(length(r_rsp_pair),100)
    rsq_s <- r_rsp_pair[select]
    rsq_resampling_all <- c(rsq_resampling_all,rsq_s)
    
    r_std_trait <- r_std_all[[trait]]
    index <- grep(paste0(factor),names(r_std_trait))
    rsq_fitting_all <- c(rsq_fitting_all,r_std_trait[[index]])
    
  }
  
  rsq_resampling_all <- as.data.frame(rsq_resampling_all)
  rsq_resampling_all$f <- "a"
  colnames(rsq_resampling_all)[1] <- "rsq"
  
  rsq_fitting_all <- as.data.frame(rsq_fitting_all)
  rsq_fitting_all$f <- "b"
  colnames(rsq_fitting_all)[1] <- "rsq"
  
  dat <- rbind(rsq_resampling_all, rsq_fitting_all)
  
  p2[[i]] <- ggplot(dat,aes(x=rsq)) + 
    
    theme_bw(base_line_size =0.3) +
    
    ggtitle(type[i]) + theme(plot.title = element_text(hjust = 0.5)) +
    
    # geom_histogram(data=subset(dat,f == 'a'),aes(fill=f,y=..density..), alpha = 0.2, binwidth = 0.05) +
    # geom_histogram(data=subset(dat,f == 'b'),aes(fill=f,y=..density..), alpha = 0.2, binwidth = 0.05) +
    # 
    # scale_fill_manual(name="group", values=c("blue", "red"),labels=c("control","real data")) +
    # 
    geom_histogram(data=subset(dat,f == 'a'),aes(y=..density..), fill="blue",alpha = 0.2, binwidth = 0.05) +
    geom_histogram(data=subset(dat,f == 'b'),aes(y=..density..), fill="red",alpha = 0.2, binwidth = 0.05) +
    
    theme(axis.title.x = element_text(size = 22),
          axis.text.x = element_text(size = 20, color = "black"),
          axis.title.y = element_text(size = 22),
          axis.text.y = element_text(size = 20, color = "black"),
          text= element_text(family="Times New Roman"))
  
  
}


all1 <- grid.arrange(p1[[1]],p1[[2]],p1[[3]],p1[[4]],p1[[5]],p1[[6]], nrow = 2)
all2 <- grid.arrange(p2[[1]],p2[[2]],p2[[3]],p2[[4]],p2[[5]],p2[[6]], nrow = 2)
```

### ECT qqplots and rsq barplots for all pairs have >=5SNPs

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE,fig.height=10, fig.width=15}


type <- c("B cell (CD19+)","Monocyte (CD14+)","Leukocyte (CD15+)","Platelet","T cell (CD4+)","T cell (CD8+)")
path <- c("4.B_cell_kegg","6.CD14_positive_monocyte_kegg","8.CD15_positive_leukocyte_kegg","16.platelet_kegg","10.T_cell_kegg","14.thymocyte_kegg")

p1 <- list()
p2 <- list()
for (i in 1:length(type)){ 
  
  file_ect_results <- paste0("/project2/xinhe/xsun/CEDAR/",path[i],"/9.pcs_40genes_rerun/2.ECT/results/ECT_results.rdata")
  load(file_ect_results)
  
  file_ect_summary <- paste0("/project2/xinhe/xsun/CEDAR/",path[i],"/9.pcs_40genes_rerun/2.ECT/results/ECT_summary_table.rdata")
  load(file_ect_summary)
  
  plot_sum <- df_sum[df_sum$num_snp_pass_cutoff >= 5,] ######
  
  plotdata <- data.frame(observed = -log10(sort(as.numeric(as.character(plot_sum$pval_ECT)))),
                         expected = -log10(ppoints(nrow(plot_sum))))
  #  
  
  p1[[i]] <- ggplot(plotdata) + theme_bw(base_line_size =0.3,) +
    geom_point(aes(expected, observed), shape = 1, size = 3, color= "steelblue") +
    geom_abline(intercept = 0, slope = 1, alpha = 0.6, color = "red") +
    ggtitle(type[i]) + theme(plot.title = element_text(hjust = 0.5)) +
    labs(x = expression(paste("Expected -log"[10],"(p-value)")),
         y = expression(paste("Observed -log"[10],"(p-value)"))) +
    
    theme(axis.title.x = element_text(size = 22),
          axis.text.x = element_text(size = 20, color = "black"),
          axis.title.y = element_text(size = 22),
          axis.text.y = element_text(size = 20, color = "black"),
          text= element_text(family="Times New Roman")) 
  
  
  #### rsq barplot (selected data)
  rsq_resampling_all <- c()
  rsq_fitting_all <- c()
  for (t in 1:nrow(plot_sum)) {
    
    trait <- plot_sum$trait[t]
    factor <- plot_sum$factor[t]
    
    r_rsp_pair <- r_rsp_all[[trait]][[factor]]
    set.seed(100)
    select <- sample(length(r_rsp_pair),100)
    rsq_s <- r_rsp_pair[select]
    rsq_resampling_all <- c(rsq_resampling_all,rsq_s)
    
    r_std_trait <- r_std_all[[trait]]
    index <- grep(paste0(factor),names(r_std_trait))
    rsq_fitting_all <- c(rsq_fitting_all,r_std_trait[[index]])
    
  }
  
  rsq_resampling_all <- as.data.frame(rsq_resampling_all)
  rsq_resampling_all$f <- "a"
  colnames(rsq_resampling_all)[1] <- "rsq"
  
  rsq_fitting_all <- as.data.frame(rsq_fitting_all)
  rsq_fitting_all$f <- "b"
  colnames(rsq_fitting_all)[1] <- "rsq"
  
  dat <- rbind(rsq_resampling_all, rsq_fitting_all)
  
  p2[[i]] <- ggplot(dat,aes(x=rsq)) + 
    
    theme_bw(base_line_size =0.3) +
    
    ggtitle(type[i]) + theme(plot.title = element_text(hjust = 0.5)) +
    
    # geom_histogram(data=subset(dat,f == 'a'),aes(fill=f,y=..density..), alpha = 0.2, binwidth = 0.05) +
    # geom_histogram(data=subset(dat,f == 'b'),aes(fill=f,y=..density..), alpha = 0.2, binwidth = 0.05) +
    # 
    # scale_fill_manual(name="group", values=c("blue", "red"),labels=c("control","real data")) +
    # 
    geom_histogram(data=subset(dat,f == 'a'),aes(y=..density..), fill="blue",alpha = 0.2, binwidth = 0.05) +
    geom_histogram(data=subset(dat,f == 'b'),aes(y=..density..), fill="red",alpha = 0.2, binwidth = 0.05) +
    
    theme(axis.title.x = element_text(size = 22),
          axis.text.x = element_text(size = 20, color = "black"),
          axis.title.y = element_text(size = 22),
          axis.text.y = element_text(size = 20, color = "black"),
          text= element_text(family="Times New Roman"))
  
  
}


all1 <- grid.arrange(p1[[1]],p1[[2]],p1[[3]],p1[[4]],p1[[5]],p1[[6]], nrow = 2)
all2 <- grid.arrange(p2[[1]],p2[[2]],p2[[3]],p2[[4]],p2[[5]],p2[[6]], nrow = 2)
```


### Top 10 pairs have ECT FDR <0.1

```{r echo=FALSE, message=FALSE, warning=FALSE}
path <- c("4.B_cell_kegg","6.CD14_positive_monocyte_kegg","8.CD15_positive_leukocyte_kegg","16.platelet_kegg","10.T_cell_kegg","14.thymocyte_kegg")
type <- c("B cell (CD19+)","Monocyte (CD14+)","Leukocyte (CD15+)","Platelet","T cell (CD4+)","T cell (CD8+)")

summary <- data.frame(matrix(NA, nrow = 1, ncol = 12))
for (i in 1:length(path)){
  folder_path <- paste0("/project2/xinhe/xsun/CEDAR/",path[i],"/9.pcs_40genes_rerun/2.ECT/")
  
  load(paste0(folder_path,"results/ECT_summary_table.rdata"))
  
  if (sum(df_sum$fdr_ECT < 0.1) > 0) {
    tmp <- df_sum[df_sum$fdr_ECT < 0.1,]
    tmp$celltype <- type[i]
    
    colnames(summary) <- colnames(tmp)
    summary <- rbind(summary,tmp)
  }
  
}
summary <- summary[-1,]
summary <- cbind(summary$celltype,summary)
summary <- summary[,-ncol(summary)]
colnames(summary)[1] <- "Celltypes"
DT::datatable(summary, options = list(pageLength =10))
```


effect size plots for the top 2 pairs

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE,fig.height=5, fig.width=6}

 load("/project2/xinhe/xsun/CEDAR/10.T_cell_kegg/9.pcs_40genes_rerun/2.ECT/results/assoc_results_harmo_2snps.rdata")
 trait <- "EUR.CD"
 factor <- "pwy124_pc1"
 fdr_cutoff <- 0.2
 dat <- assoc_harmo_all[[trait]][[factor]]
 snp_select <- dat[dat$fdr < fdr_cutoff,]

 fit <- lm(snp_select$beta.outcome ~ 0 + snp_select$beta.exposure)

 eff <- ggplot(snp_select, aes(x=beta.exposure, y=beta.outcome, color=chr)) +
 geom_point() +
 geom_errorbar(aes(ymin=beta.outcome-se.outcome, ymax=beta.outcome+se.outcome, color=chr), width=0.01 ,alpha=0.6) +
 theme_bw(base_line_size =0.3) +
 theme(axis.title.x = element_text(size = 12),
       axis.text.x = element_text(size = 10, color = "black"),
       axis.title.y = element_text(size = 12),
       axis.text.y = element_text(size = 10, color = "black"),
       text= element_text(family="Times New Roman")) +
 geom_smooth(method = lm, se = F, color = "grey", size = 0.5) +
 geom_label(aes(x = -0.25, y = 0.07), hjust = 0,
            label = paste("rsq = ",signif(summary(fit)$r.squared, 3),
                          " \nECT p-value =5.0e-07"),
                                  #" \np-value =",signif(summary(fit)$coef[1,4], 3)),
            color = "black",
            family = "Times New Roman",
            size = 3,
            label.size = NA,
            alpha = 0) +
   ggtitle(paste0(factor,"-", trait)) + theme(plot.title = element_text(hjust = 0.5))
 
eff
```


```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE,fig.height=5, fig.width=6}

 load("/project2/xinhe/xsun/CEDAR/14.thymocyte_kegg/9.pcs_40genes_rerun/2.ECT/results/assoc_results_harmo_2snps.rdata")
 trait <- "T1D"
 factor <- "pwy122_pc5"
 fdr_cutoff <- 0.2
 dat <- assoc_harmo_all[[trait]][[factor]]
 snp_select <- dat[dat$fdr < fdr_cutoff,]

 fit <- lm(snp_select$beta.outcome ~ 0 + snp_select$beta.exposure)

 eff <- ggplot(snp_select, aes(x=beta.exposure, y=beta.outcome, color=chr)) +
 geom_point() +
 geom_errorbar(aes(ymin=beta.outcome-se.outcome, ymax=beta.outcome+se.outcome, color=chr), width=0.01 ,alpha=0.6) +
 theme_bw(base_line_size =0.3) +
 theme(axis.title.x = element_text(size = 12),
       axis.text.x = element_text(size = 10, color = "black"),
       axis.title.y = element_text(size = 12),
       axis.text.y = element_text(size = 10, color = "black"),
       text= element_text(family="Times New Roman")) +
 geom_smooth(method = lm, se = F, color = "grey", size = 0.5) +
 geom_label(aes(x = -0.18, y = 0.07), hjust = 0,
            label = paste("rsq = ",signif(summary(fit)$r.squared, 3),
                          " \nECT p-value =5.0e-07"),
                                  #" \np-value =",signif(summary(fit)$coef[1,4], 3)),
            color = "black",
            family = "Times New Roman",
            size = 3,
            label.size = NA,
            alpha = 0) +
   ggtitle(paste0(factor,"-", trait)) + theme(plot.title = element_text(hjust = 0.5))
 
eff
```

### Detailed results 

#### B cell (CD19+)

```{r echo=FALSE, message=FALSE, warning=FALSE}
load("/project2/xinhe/xsun/CEDAR/4.B_cell_kegg/9.pcs_40genes_rerun/2.ECT/results/ECT_summary_table.rdata")
DT::datatable(df_sum, options = list(pageLength =5))
```

Compute ECT FDR only for pairs have >=5 SNPs



```{r echo=FALSE, message=FALSE, warning=FALSE}
df_sum_15 <- df_sum[df_sum$num_snp_pass_cutoff >= 5,]
df_sum_15$ECT_fdr_5snp <- p.adjust(df_sum_15$pval_ECT, method = "fdr")
DT::datatable(df_sum_15, options = list(pageLength =5))
```




#### Monocyte (CD14+)

```{r echo=FALSE, message=FALSE, warning=FALSE}
load("/project2/xinhe/xsun/CEDAR/6.CD14_positive_monocyte_kegg/9.pcs_40genes_rerun/2.ECT/results/ECT_summary_table.rdata")
DT::datatable(df_sum, options = list(pageLength =5))
```

Compute ECT FDR only for pairs have >=5 SNPs


```{r echo=FALSE, message=FALSE, warning=FALSE}
df_sum_15 <- df_sum[df_sum$num_snp_pass_cutoff >= 5,]
df_sum_15$ECT_fdr_5snp <- p.adjust(df_sum_15$pval_ECT, method = "fdr")
DT::datatable(df_sum_15, options = list(pageLength =5))
```
#### Leukocyte (CD15+)

```{r echo=FALSE, message=FALSE, warning=FALSE}
load("/project2/xinhe/xsun/CEDAR/8.CD15_positive_leukocyte_kegg/9.pcs_40genes_rerun/2.ECT/results/ECT_summary_table.rdata")
DT::datatable(df_sum, options = list(pageLength =5))
```

Compute ECT FDR only for pairs have >=5 SNPs


```{r echo=FALSE, message=FALSE, warning=FALSE}
df_sum_15 <- df_sum[df_sum$num_snp_pass_cutoff >= 5,]
df_sum_15$ECT_fdr_5snp <- p.adjust(df_sum_15$pval_ECT, method = "fdr")
DT::datatable(df_sum_15, options = list(pageLength =5))
```

#### Platelet

```{r echo=FALSE, message=FALSE, warning=FALSE}
load("/project2/xinhe/xsun/CEDAR/16.platelet_kegg/9.pcs_40genes_rerun/2.ECT/results/ECT_summary_table.rdata")
DT::datatable(df_sum, options = list(pageLength =5))
```

Compute ECT FDR only for pairs have >=5 SNPs


```{r echo=FALSE, message=FALSE, warning=FALSE}
df_sum_15 <- df_sum[df_sum$num_snp_pass_cutoff >= 5,]
df_sum_15$ECT_fdr_5snp <- p.adjust(df_sum_15$pval_ECT, method = "fdr")
DT::datatable(df_sum_15, options = list(pageLength =5))
```



#### T cell (CD4+)

```{r echo=FALSE, message=FALSE, warning=FALSE}
load("/project2/xinhe/xsun/CEDAR/10.T_cell_kegg/9.pcs_40genes_rerun/2.ECT/results/ECT_summary_table.rdata")
DT::datatable(df_sum, options = list(pageLength =5))
```

Compute ECT FDR only for pairs have >=5 SNPs


```{r echo=FALSE, message=FALSE, warning=FALSE}
df_sum_15 <- df_sum[df_sum$num_snp_pass_cutoff >= 5,]
df_sum_15$ECT_fdr_5snp <- p.adjust(df_sum_15$pval_ECT, method = "fdr")
DT::datatable(df_sum_15, options = list(pageLength =5))
```


#### T cell (CD8+)

```{r echo=FALSE, message=FALSE, warning=FALSE}
load("/project2/xinhe/xsun/CEDAR/14.thymocyte_kegg/9.pcs_40genes_rerun/2.ECT/results/ECT_summary_table.rdata")
DT::datatable(df_sum, options = list(pageLength =5))
```

Compute ECT FDR only for pairs have >=5 SNPs

```{r echo=FALSE, message=FALSE, warning=FALSE}
df_sum_15 <- df_sum[df_sum$num_snp_pass_cutoff >= 5,]
df_sum_15$ECT_fdr_5snp <- p.adjust(df_sum_15$pval_ECT, method = "fdr")
DT::datatable(df_sum_15, options = list(pageLength =5))
```



