---
title: "ldsc"
author: "XSun"
date: "2020-06-25"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction

I adopted LD score regression to estimate SNP-based heritability and partation the heritability into seperate categories. The software I used here is [ldsc](https://github.com/bulik/ldsc).


## Material and Methods

'ldsc' takes GWAS summary statistics, LD scores, regression weights, and allele frequenciesas inputs. Regression weights, allele frequenciesas and 53 categories of LD scroes can be downloaded [here](https://data.broadinstitute.org/alkesgroup/LDSCORE/). We need to prepare our own GWAS summary statistics and LD scores.


### LD scores

I used 13 histone marks (H3K27ac,DNase,H3K27me3,H3K36me3,H3K4me1,H3K4me3,H3K9me3,H3K9ac,H2A,H3K4me2,H3K79me2,H4K20me1,H3K9me1) and open chromatin regions in immune cells to compute partitioned LD scores. The marks and cell types are listed below.


```{r echo=FALSE}
aa <- read.table("/project2/xinhe/xsun/website/factor_analysis/output/annot_list_2col.txt", header = F)
colnames(aa) <- c("Marks","Cell types")

DT::datatable(aa, options = list(pageLength =10))
```

Apart from these annotations, I also used annotation

### Summary Statistics 

The software 'ldsc' uses GWAS summary statistics as inputs. The GWAS summary statistics we adopted here came from the association tests for factors ~ variants + covariants(using training data, 536 samples). I performed the test for both maf>0.05 variants and maf>0.01 variants. The covariants here are 1.top ten pcs from genotype matrix 2.factors whose heritability equal to 0.000001 and p_val_test_l is greater than 0.5.

I converted the GWAS results into following format(taking lv73, the one has the highest heritability in plier factors_allpathway as an example).


```{r eval=FALSE, echo=T}
head(sum_plier_allpathway_lv73.txt)
SNP N Z A1 A2
rs112409036 535 -0.160132099643991 T A
rs11166389 533 0.00275588338106556 A G
rs116740877 536 -0.0102126413071695 A T
rs7535103 536 0.0275829957516441 T C
rs2149190 523 0.0167134986922463 C G
```
      

The 'SNP' IDs here are from dbSNP 151, I downloaded the correspondence between dbSNP 151 and GTEx variant ID from the [GTEx portal](https://storage.googleapis.com/gtex_analysis_v8/reference/GTEx_Analysis_2017-06-05_v8_WholeGenomeSeq_838Indiv_Analysis_Freeze.lookup_table.txt.gz)

The 'N' here stands for the the number of non-missing samples for each variants.

'Z' is the z-score for each variants. I computed it from p-value and effect size. 
Z=sign(Effect Size)×|phi^(-1)(p/2)|.  phi^(-1) is the inverse cumulative distribution function of the normal distribution.

A1 is the alternative allele and A2 is the reference allele.
